---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# UniTED: A Unified Time Series Event Detection Repository powered by DAL

<!-- badges: start -->
![GitHub Repo stars](https://github.com/cefet-rj-dal/united)
<!-- badges: end -->


The Unified Event Detection Dataset (UniTED), powered by DAL, is a curated
repository of time series datasets for the development and evaluation of event
and anomaly detection methods. It primarily includes labeled datasets to enable
consistent benchmarking and reproducible experiments.

This repository contains:
- The R package that exposes dataset objects (see `?data` topics and `man/`)
- The data files shipped for package use (`data/*.RData`)
- The ETL pipelines used to build those data files (`etl/`), separated into:
  - `source/`: original or labeled inputs used by ETL
  - `intermediate/`: zipped files and grouped lists produced during processing
  - final artifacts are saved under the repo root `data/`

We recommend accessing datasets via this package to ensure standardized
structures and metadata.

---

## Dataset Description

- Full dataset descriptions: <https://github.com/cefet-rj-dal/united/wiki>
- Example notebooks: `Rmd/examples` and rendered versions in `examples/`
- ETL documentation: one README per dataset under `etl/<dataset>/README.md`


---

## Examples

Usage examples are organized by dataset: <https://github.com/cefet-rj-dal/united/tree/main/examples/>


```{r example}
library(united)
# Load mini dataset and fetch full version
data(A1Benchmark)
A1Benchmark <- loadfulldata(A1Benchmark)
```

---

## Installation

You can install the published version of UniTED package from GitHub [https://github.com/cefet-rj-dal/united](https://github.com/cefet-rj-dal/united) with:

``` r
timeout <- options()$timeout
options(timeout=1200)
devtools::install_github("cefet-rj-dal/united", upgrade="never")
options(timeout=timeout)
```


## Bugs and new features request
<https://github.com/cefet-rj-dal/united/issues>

---

## ETL Generation

The `etl/` folder hosts dataset‑specific ETL pipelines that produce the final
RData files included in the package. Each dataset folder (e.g., `etl/numenta`,
`etl/ucr_anomaly`, `etl/yahoo`, `etl/3W`, `etl/mit-bih`, `etl/gecco`) contains:

- `source/`: the labeled inputs used by builder scripts
- `intermediate/`: zipped/raw conversions and grouped list artifacts
- builder scripts: numbered `*.R` files that specify the ETL order and outputs
- a README with origin links, field descriptions, and script references

Final artifacts are placed in the root `data/` directory and documented in
`R/united.R`. To reproduce data, run the numbered scripts inside `etl/<dataset>/`.

---

## References

- Lima, J., Castro, H., Oliveira, L., Paixão, E., Baroni, L., Salles, R., 
  Vargas, R., Ogasawara, E. UniTED: A Unified Time Series 
  Event Detection Repository. Brazilian e-Science Workshop (BreSci). 
  In: Brazilian e-Science Workshop (BreSci). SBC, DOI: 10.5753/bresci.2025.247972. 
- Chandola, V., Banerjee, A., & Kumar, V. (2009). Anomaly detection: A survey.
  ACM Computing Surveys, 41(3), 1–58.
- Truong, C., Oudre, L., & Vayatis, N. (2020). Selective review of change point
  detection methods. Signal Processing, 167, 107299.
- Lavin, A., & Ahmad, S. (2015). Evaluating real-time anomaly detection algorithms
  — the Numenta Anomaly Benchmark. 2015 IEEE 14th ICMLA.

