---
title: "3W Dataset - Oil wells"
output: html_notebook
---
The first realistic and public dataset with rare undesirable real events in oil wells.

* Multivariate series with labeled change points
* Recommended use: multivariate or univariate CPD detection

Source: https://archive.ics.uci.edu/ml/datasets/3W+dataset


## Load series
```{r}
library(dalevents)
library(daltoolbox)
library(harbinger)
```

Selecting a well as example

```{r}
## Load series ----------------------
data(oil_3w_Type_1)
```


```{r}
#Selecting
data <- oil_3w_Type_1[[1]]
```


## Univariate series selection
Select the desired variable directly from preprocessed data.

```{r}
series <- data$p_tpt
plot(as.ts(series))
```

## Event detection experiment


Detection steps
```{r}
#Establishing arima method
model <- hcp_binseg()
```


```{r}
#Fitting the model
model <- fit(model, series)
```


```{r}
#Making detections
detection <- detect(model, series)
```


## Results analysis


```{r}
#Filtering detected events
print(detection |> dplyr::filter(event==TRUE))
```

Visual analysis
```{r}
#Ploting the results
grf <- har_plot(model, series, detection, data$event)
plot(grf)
```

Evaluate metrics
```{r}
#Evaluating the detection metrics
ev <- evaluate(model, detection$event, data$event)
print(ev$confMatrix)
```

```{r}
ev$accuracy
ev$F1
```

### SoftEd Evaluation
To analyze the results considering temporal tolerance, softED smoothed metrics can be used, as performed below.

```{r}
ev_soft <- evaluate(har_eval_soft(sw=90), detection$event, data$event)
print(ev_soft$confMatrix)
```

```{r}
ev_soft$accuracy
ev_soft$F1
```
