---
title: "3W Dataset - Oil wells"
output: html_notebook
---
The first realistic and public dataset with rare undesirable real events in oil wells.

* Multivariate series with labeled change points
* Recommended use: multivariate or univariate CPD detection

Source: https://archive.ics.uci.edu/ml/datasets/3W+dataset


## Load series
```{r}
library(dalevents)
library(daltoolbox)
library(harbinger)


## Load series ----------------------
data(oil_3w_Type_1)
```

Selecting a well as example

```{r}
#Selecting
data <- oil_3w_Type_1$Type_1$`WELL-00001_20140124213136`
summary(data)
```

## Preprocessing

* Remove NA values

```{r}
#Removing empty variables
data$T_JUS_CKGL <- NULL

plot(as.ts(data))
```

* Normalize data

```{r}
#data <- na.omit(data)
ts <- data[,1:7]

preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

head(ts)
```
```{r}
plot(as.ts(ts))
```

## Labels adjust

It is essential to understand the dataset, including data types and intended use, to ensure accurate interpretation of labels. In the 3W dataset, the classes represent transitional stages in behavior observed during oil well drilling monitoring. For event detection purposes, the labels are being adjusted to treat each stage transition as a distinct event.


```{r}

labels <- as.data.frame(data$class)
names(labels) <- "class"
plot(as.ts(labels$class))

labels$cpd <- 0
head(labels)

#Initial CP
cp = FALSE
for (i in 1:nrow(labels)){
  if (cp == FALSE){
    if(!is.na(labels$class[i]) && labels$class[i] != 0){
      print("Ponto de mudança rotulado em:")
      print(i)
      labels$cpd[i] <- 1
      cp = i
    }
  }
}

#Final CP
i = cp+1
cp = FALSE
print(i)


for (i in 1018:nrow(labels)){
  if (cp == FALSE){
    if(!is.na(labels$class[i]) && labels$class[i] != 101){
      print("Ponto de mudança rotulado em:")
      print(i)
      labels$cpd[i] <- 1
      cp = i
    }
  }
}

#Count of CP
sum(labels$cpd)

plot(as.ts(labels$cpd),
     main = "Change Points")
```


## Univariate series selection
Select the desired variable directly from preprocessed data.

```{r}
series <- ts$P_MON_CKP
plot(as.ts(series))
```

## Event detection experiment

Creating a data frame to organize experiment results.

```{r}
#Experiments results organization
experiment <- data.frame(method="hanr_arima",
                         dataset="3W",
                         series="Type_1_WELL_00001_20140124213136_P_MON_CKP",
                         elapsed_time_fit=0,
                         elapsed_time_detection=0,
                         accuracy=0,
                         precision=0,
                         recall=0,
                         F1=0)

head(experiment)
```
Detection steps
```{r}
#Establishing arima method
model <- hanr_arima()
```


```{r}
#Fitting the model
s <- Sys.time()
model <- fit(model, series)
t_fit <- Sys.time()-s
```


```{r}
#Making detections
s <- Sys.time()
detection <- detect(model, series)
t_det <- Sys.time()-s
```


## Results analysis


```{r}
#Filtering detected events
print(detection |> dplyr::filter(event==TRUE))
```

Visual analysis
```{r}
#Ploting the results
grf <- har_plot(model, series, detection, labels$cpd)
plot(grf)
```

Evaluate metrics
```{r}
#Evaluating the detection metrics
ev <- evaluate(model, detection$event, labels$cpd)
print(ev$confMatrix)
```

Recording experiment results
```{r}
#Experiment update
#Time
experiment$elapsed_time_fit[1] <- as.numeric(t_fit)*60
experiment$elapsed_time_detection[1] <- as.numeric(t_det)*60

#Metrics
experiment$accuracy[1] <- ev$accuracy
experiment$precision[1] <- ev$precision
experiment$recall[1] <- ev$recall
experiment$F1[1] <- ev$F1

print(experiment)
```

### SoftEd Evaluation
To analyze the results considering temporal tolerance, softED smoothed metrics can be used, as performed below.

```{r}
ev_soft <- evaluate(har_eval_soft(sw=90), detection$event, as.logical(labels$cpd))
print(ev_soft$confMatrix)
```

```{r}
ev_soft$accuracy
ev_soft$F1
```

### Add new series to experiment 
Repeat detection steps

* Add new rows to the `experiment` data frame
* Run detection steps (create model, fit and detect) for the new series
* Update `experiment` with the new series result
  * Repeated steps for didactic purposes
  * WARNING: In real experimental situations, variable selection and repetition of detection steps should be encapsulated in a loop or function

### Experiment record
The retain experiments data record results

* Save detection results: Save `detection` object after detection of each series
* Save experiment: Save `experiment` object after finishing the complete experiment
