---
title: "Gecco Challenge Dataset - Water Quality"
output: html_notebook
---
Data collection for water quality monitoring

* Multivariate series with labeled anomalies.
* Recommended use: multivariate or univariate event detection

Source: https://www.spotseven.de/gecco/gecco-challenge


## Load series
```{r}
library(dalevents)
library(daltoolbox)
library(harbinger)


## Load series ----------------------
data(gecco)
```

Gecco recommended sample: One day with anomalies

```{r}
data <- gecco$gecco
data <- data[16500:18000,]
plot(as.ts(data))
```



## Univariate series selection
Select the desired variable directly from preprocessed data.

```{r}
series <- data$ph
plot(as.ts(series))
```

## Event detection experiment

Detection steps
```{r}
#Establishing arima method
model <- hanr_arima()
```


```{r}
#Fitting the model
s <- Sys.time()
model <- fit(model, series)
t_fit <- Sys.time()-s
```


```{r}
#Making detections
s <- Sys.time()
detection <- detect(model, series)
t_det <- Sys.time()-s
```


## Results analysis


```{r}
#Filtering detected events
print(detection |> dplyr::filter(event==TRUE))
```

Visual analysis
```{r}
#Ploting the results
grf <- har_plot(model, series, detection, data$event)
plot(grf)
```

Evaluate metrics
```{r}
#Evaluating the detection metrics
ev <- evaluate(model, detection$event, data$event)
print(ev$confMatrix)
```
