---
title: "RARE: a labeled dataset for cloud-native memory anomalies"
output: html_notebook
---
Cloud-native memory anomalies

* Multivariate series with labeled anomalies
* Recommended use: multivariate or univariate event detection
* WARNING: Example under construction - This dataset is under our analysis for better organization and suggested use.

Source: https://dl.acm.org/doi/10.1145/3416505.3423560


## Load series
```{r}
library(dalevents)
library(daltoolbox)
library(harbinger)


## Load series ----------------------
data(rare)
```


## Univariate series selection
Select the desired variable.

```{r}
series <- rare[2]
plot(as.ts(series))
```

## Labels
```{r}
series$event <- rare$event
names(series) <- c("value", "event")


plot(as.ts(series))
```


## Event detection experiment

Detection steps
```{r}
#Establishing arima method
model <- hanr_arima()
```


```{r}
#Fitting the model
s <- Sys.time()
model <- fit(model, series$value)
t_fit <- Sys.time()-s
```


```{r}
#Making detections
s <- Sys.time()
detection <- detect(model, series$value)
t_det <- Sys.time()-s
```


## Results analysis


```{r}
#Filtering detected events
print(detection |> dplyr::filter(event==TRUE))
```

Visual analysis
```{r}
#Ploting the results
grf <- har_plot(model, series$value, detection, series$event)
plot(grf)
```

Evaluate metrics
```{r}
#Evaluating the detection metrics
ev <- evaluate(model, detection$event, series$event)
print(ev$confMatrix)
```


### SoftEd Evaluation
To analyze the results considering temporal tolerance, softED smoothed metrics can be used, as performed below.

```{r}
ev_soft <- evaluate(har_eval_soft(sw=90), detection$event, as.logical(series$event))
print(ev_soft$confMatrix)
```

```{r}
ev_soft$accuracy
ev_soft$F1
```
